\section{Problem Setup}

\subsection{State and Forecast Target}

Each sample is anchored at label time $t_0$. The observed history contains boxes up to and including $t_0$, and the target is the sequence of future boxes after $t_0$. A box is represented in normalized center-size coordinates
\begin{equation}
  \bbox_t = [c^x_t, c^y_t, w_t, h_t]^\top \in [0,1]^4,
\end{equation}
where coordinates are normalized by the frame size. The model state augments the box with first-order velocity:
\begin{equation}
  \state_t =
  [c^x_t, c^y_t, w_t, h_t, v^x_t, v^y_t, v^w_t, v^h_t]^\top.
\end{equation}

The main configuration uses a $400$ ms history window and an $800$ ms forecast window, with all selected tracks required to contain at least $1.2$ s of usable labeled duration. Samples are generated from track-aligned FRED labels in \texttt{Event\_YOLO} and \texttt{cleaned\_tracks.txt}. Event inputs use trailing render windows to avoid future leakage. We use a random folder-level holdout from the training split for validation during development, with \texttt{train\_eval\_split.enabled=true} and a held-out fraction of $0.2$ unless otherwise stated.

\subsection{Inputs}

Image-conditioned runs use one configured rendered or dataset-native representation at a time. To reduce direct position leakage and keep training feasible, the main image-conditioned experiments use fixed cutouts around the final history box. In the current implementation these cutouts are $64\times64$ at the resized input resolution, corresponding to $128\times128$ pixels in the original FRED frame resolution. All image inputs are resized to $640\times360$ before cutout extraction. For time-space projections, the temporal dimension is fixed to $64$ bins.

The evaluated input families are:
\begin{itemize}
  \item \textbf{CSTR2/CSTR3}: compact event representations rendered over the trailing event window;
  \item \textbf{XT/YT}: time-space projections chosen to avoid direct mean-position channels where possible;
  \item \textbf{event frames}: dataset-native frames from \texttt{Event/Frames} when available;
  \item \textbf{RGB}: \texttt{rgb} or \texttt{padded\_rgb} frames aligned to the anchor time when used as a single image input; and
  \item \textbf{box and filter-state features}: recent normalized boxes and optimized Kalman filter state features.
\end{itemize}

For XT inputs, spatial cutouts affect the $x$ axis only while preserving the temporal axis. For YT inputs, cutouts affect the $y$ axis only. Other image-like inputs use both spatial axes.

To limit explicit absolute-position information, the residual model uses relative box-history features in the main experiments. When filtered center positions are used as learned features, we optionally map normalized coordinates from $[0,1]$ to $[-1,1]$ so that the frame center is zero.

\subsection{Metrics}

Predictions are evaluated against future boxes using:
\begin{align}
  \mathrm{ADE}_{center} &= \frac{1}{T}\sum_{k=1}^{T}
  \left\|\hat{\mathbf{c}}_{t_k} - \mathbf{c}_{t_k}\right\|_2, \\
  \mathrm{FDE}_{center} &= 
  \left\|\hat{\mathbf{c}}_{t_T} - \mathbf{c}_{t_T}\right\|_2,
\end{align}
reported in pixels. We also report bounding-box ADE/FDE in pixels and mean IoU over the forecast horizon.

\todo{State exact train, validation, and test folder counts once the final split policy is frozen.}
